---
title: "RStudio Addins http://rstudio.github.io/rstudioaddins/"
author: "Andy Lathrop"
date: "March 24, 2016"
output: 
  html_document:
    theme: united
    toc: TRUE
    toc_float: TRUE
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About
#### RStudio addins provide a mechanism for executing R functions interactively from within the RStudio IDE—either through keyboard shortcuts, or through the Addins menu.
#### An addin can be as simple as a function that inserts a commonly used snippet of text, and as complex as a Shiny application that accepts input from the user, and later mutates a document open in RStudio. The sky is the limit!


```{r}
# ensure package 'pacman' for package management is installed
if (!require("pacman")) install.packages("pacman")
library(pacman)

# install and load required packages
pacman::p_load(devtools,
               rstudioapi,
               addinexamples, 
               update=TRUE)
pacman::p_loaded()
```


# R addin development
## Will need to create an R package
#### more info at <http://bit.ly/22y7OMc>
```{r}
insertMyAddin <- function() {
  rstudioapi::insertText("getOption(\"repos\")")
}
```

