---
title:  "rmarkdown template"
author: "Author: Andy Lathrop"
date:   "Date: `r Sys.Date()`"
output: 
  html_document:
    theme: united
    toc: TRUE
    toc_float: TRUE
    code_folding: show
---
***
##### Use the floating **table of contents** on the left for navigation, and **gray boxes near code chunks** to show/hide

Admin notes {.tabset}
=======================
***
**rmarkdown** package required
------------------------------
#### This is an R markdown (.Rmd) document and requires the **'rmarkdown'** package
#### Use the tabs in this section for installation information
#### more info at <http://rmarkdown.rstudio.com/>
***

**rmarkdown** installation
--------------------------
#### If you are working within RStudio then you can simply install the current release of RStudio (both the rmarkdown package and pandoc are included). <https://www.rstudio.com/products/rstudio/download/preview/>
#### The code below will install the 'rmarkdown' package if needed
```{r rmarkdown install}
if (!require("rmarkdown")) install.packages("rmarkdown", repos='https://cran.revolutionanalytics.com')
library(rmarkdown)
```
***
#### **rmarkdown** global setup options  
#####Setup options: echo means R code will be displayed | cache means results will be stored for future knits
```{r setup, include=TRUE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE)
```
***
package management 
------------------
#### This section manages all packages required for this script
##### The 'p_load()' function from the pacman package is the base equivalent of 'install.packages()' AND 'library()'
```{r package management}
# ensure package 'pacman' for package management is installed
if (!require("pacman")) install.packages("pacman")
library(pacman)

# add any required packages to the p_load() function parameters 
pacman::p_load(rmarkdown, 
               e1071, # for support vector machines 
               update=TRUE)
pacman::p_loaded() # check which packages are loaded
```

***
about
-----
##### Built with **Microsoft R Open** version `r Revo.version`. More info at <https://mran.revolutionanalytics.com/open/>
##### The current CRAN repository for this R session is `r getOption("repos")`
##### The current default MRAN repository can be found at <https://mran.revolutionanalytics.com/timemachine/>
***